---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
df <- read.csv("./data/activity.csv", stringsAsFactors=FALSE)
df$date <- as.Date(df$date, "%Y-%m-%d")
df$interval <- as.numeric(df$interval)
```

## What is mean total number of steps taken per day?
```{r}
steps <- tapply(df$steps, df$date, sum)
hist(steps)

mean_steps <- mean(steps, na.rm = TRUE)
median_steps <- median(steps, na.rm = TRUE)
```
The mean and median of the total number of steps taken per day are `r mean_steps` and `r median_steps` respectively.

## What is the average daily activity pattern?
```{r}
avgs <- tapply(df$steps, df$interval, mean, na.rm = TRUE)
plot(names(avgs), avgs, type = "l", xlab = "Interval")
max_avg <- names(avgs)[which.max(avgs)]
```
Interval `r max_avg`, on average across all the days in the dataset, contains the maximum number of steps

## Imputing missing values
```{r}
missing <- is.na(df$steps)
n.missing <- sum(missing)
not_missing <- df$steps[!missing]
# randomly assign non-missing values
df$steps[missing] <- sample(not_missing, n.missing, replace=TRUE)

steps <- tapply(df$steps, df$date, sum)
hist(steps)

mean_steps <- mean(steps, na.rm = TRUE)
median_steps <- median(steps, na.rm = TRUE)
```
The total number of missing values in the dataset is `r n.missing`.
The mean and median of the total number of steps taken per day are `r mean_steps` and `r median_steps` respectively.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
ends <- weekdays(df$date) == "Saturday" | weekdays(df$date) == "Saturday"
df$weekday <- as.factor(ifelse(ends, "weekend", "weekday"))

weekends <- df[ends,]
weekdays <- df[!ends,]
weekend_avg <- tapply(weekends$steps, weekends$interval, mean)
weekday_avg <- tapply(weekdays$steps, weekdays$interval, mean)
range <- range(weekend_avg, weekday_avg)

par(mfrow = c(1,2))
plot(names(weekend_avg), weekend_avg, type = "l", ylab = "Number of steps", xlab = "Interval on weekends", ylim = range)
plot(names(weekday_avg), weekday_avg, type = "l", ylab = "Number of steps", xlab = "Interval on weekdays", ylim = range)
```

